---
title: "Module 4: Assignment 3"
output: pdf_document
date: "2023-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Practicing for the Final Project

This is meant to help you prepare for the final project, which I will be handing out next week. The final project will have about the same amount of guidance (or maybe even a bit more) than I've given you here. If you can work your way through these practice problems, you'll be in good shape!

I'm expecting you to be able to filter and summarize the data in ways you need, choose the appropriate visualization, choose the appropriate analysis, and correctly interpret the analysis for the question I've asked you.

Some important notes:

-   For both this and the final project, we are going to use the `palmerpenguins` dataset, which we've used before! You can learn more about it [here](https://allisonhorst.github.io/palmerpenguins/).

-   I will not be giving you an answer key for your final project since part of your grade for the assignment is to choose the correct visualizations and analyses.

-   For every plot you make, be sure to clarify and clean up the labels and add a theme.

-   Because we are working through this together in class, I will be grading this assignment on completion only! That said, make sure you actually answer ***every question*** that's in here for full credit...there are quite a few!

## Set-Up

Let's load our packages and get started!

```{r}
library(tidyverse)
library(palmerpenguins)
```

We will be using the `penguins` data frame. It exists as part of the `palmerpenguins` package, but if you want it to show up in your environment, run the following code chunk.

```{r}
penguins <- penguins
```

## Problem Set 1

For this problem set, we want to know if there is a significant difference between the flipper lengths of the species on Dream island.

First things first, let's make a data frame with only penguins on Dream island. Name this dataframe `dream`. *We will want to use this data frame for the rest of this problem set.*

```{r}

```

Based on our question, identify which variable is independent and which is dependent. Also identify if they are continuous or categorical.

Using the `dream` data frame that you created above, calculate the minimum, maximum, and mean of the flipper lengths for each species

```{r}

```

Choose an effective visualization method for this data and plot the data using `ggplot2`.

```{r}

```

Write out the null and alternative hypotheses.

**Null:**

**Alternative:**

Run the appropriate statistical test.

```{r}

```

Interpret the results of your statistical test:

-   What is the p-value?
-   Is the p-value above or below 0.05?
-   What does your answer to the question above mean?
-   Should we reject or fail to reject the null hypothesis?

Should we run pairwise comparisons? If yes, do so below and interpret:

```{r}

```

## Problem Set 2

For this problem set, we want to know if there is a relationship between flipper length and bill length amongst all penguins (we aren't going to worry about species right now).

Are our variables of interest continuous and/or categorical?

Take note that in this example, there is no dependent or independent variable per say. We don't have any reason to think that flipper length influences bill length or vice versa. We just want to determine if there is a relationship or not.

That said, treat flipper length as the *independent* variable (x-axis) and bill length as the *dependent* variable (y-axis).

First, calculate the mean and standard deviation for both bill length and flipper length.

```{r}

```

Choose the best way to visualize the relationship between these two variables and create that plot.

```{r}

```

To do the rest of this problem set, the easiest way is to create a new data frame with no NA values. Run the code chunk below and use that new data frame for the rest of the problem set.

```{r}
penguins_noNA <- penguins %>% 
  filter(!is.na(bill_length_mm),
         !is.na(flipper_length_mm))
```

Using the `penguins_noNA` data frame, calculate the correlation coefficient and the `r^2` value.

-   According to the correlation coefficient, is the relationship positive, negative or is there no relationship?
-   According to the `r^2` value, how much of the variation in this relationship is explained? Remember, we usually multiply the `r^2` by 100 to represent this as a percentage.

```{r}

```

Run the appropriate model for this data.

```{r}

```

Using both variables and numbers from the summary above, write out the equation of the line

*Answer:*

Interpret the results of your statistical test:

-   What is the p-value?
-   Is the p-value above or below 0.05?
-   What does your answer to the question above mean?

Should we run pairwise comparisons? If yes, do so below and interpret:

```{r}

```

### Bonus

Let's add in the species variable into our analysis! Keep using the `penguins_noNA` data frame.

First, plot the data, this time including species as a variable in the plot.

```{r}

```

Run the model again, this time including species and the species interaction with "independent" variable.

```{r}

```

Interpret the results from the model above. Focus on the p-values for the independent variables and/or interaction terms, not the overall model.

*Answer:*

## Problem Set 3

For this problem set, we want to know if there is a difference in the body mass between the penguin species.

Based on our question, which variable is independent and which is dependent? Which is continuous and/or which is categorical?

First, calculate the minimum, maximum, and mean of the body mass for each species. Call this data frame `body_mass`.

```{r}

```

Convert all the columns in the `body_mass` data frame from grams to kilograms. (Remember that you divide grams by 1000 to get kilograms)

```{r}

```

Back to the question at hand...choose an effective visualization method for this data (you will want to use the original `penguins` dataframe). Practice adding lines that represent to mean values in the plot.

```{r}

```

Write out the null and alternative hypotheses.

**Null:**

**Alternative:**

Run the appropriate statistical test.

```{r}

```

Interpret the results of your statistical test:

-   What is the p-value?
-   Is the p-value above or below 0.05?
-   What does your answer to the question above mean?
-   Should we reject or fail to reject the null hypothesis?

Should we run pairwise comparisons? If yes, do so below and interpret:

```{r}

```
