---
title: "Module 1, Assignment 3: <br> Getting to Know Your Home"
author: "Ellen Bledsoe"     # <- your name here!
date: "`r Sys.Date()`"      # <- uses the current date when rendered
output: 
  github_document: default
  html_document: default
  pdf_document: default
---

# Assignment Description

### Purpose

### Task

### Criteria for Success




Assignment 1-3 - Getting to know your new home

For this final assignment for module 1, you'll be working with another real-world data set, a collection of data from climate stations scattered across antarctica. The file can be downloaded here (https://tinyurl.com/uwf9qzk). 

Your goal for this assignment is to use all of the programming skills you've acquired so far to provide some insight into this complex and large data set. 

The assignment is to write an R script that is properly formatted with best-coding practices we've discussed that accomplishes the following:  

1) Loads in the tidyverse packages to the workspace, and reads in the data  
2) Performs a quality control check on the data and notes the number of columns and rows in commented notes  
3) How many unique stations are there reported here? (Hint: look up the help file for the unique() function)  
4) Write a chunk of code that calculates the mean temperature for the month of January for each station. You can use base R or tidyverse, but it will probably be easier in tidyverse. Hint: think about using the pipe, group_by(), filter() and summarize().   
5) Which station had the coldest average temperature in January (hint: look up the arrange() function)  

Upload the script to D2L when finished.   

# Assignment Questions

1. Load in tidyverse and read in the data. Remember to use read_csv().
```{r}
# packages
library(tidyverse)

# loading in the data
df <- read_csv("../data/aggregated_station_data.csv")
```

2. Perform a quality control check on the data and note the number of columns and rows in commented notes
```{r} 
# QC
glimpse(df)

#13 columns and 138,920 rows 
```


3. How many unique stations are there reported here? (Hint: look up the help file for the unique() function)
```{r}
# Unique Stations
unique(df$station_id)
```

4. Write a chunk of code that calculates the mean temperature for the month of January for each station. You can use base R or tidyverse, but it will probably be easier in tidyverse. Hint: think about using the pipe, group_by(), filter() and summarize(). 
```{r}
# Mean for Jan for each station
df %>%
  filter(month == 1) %>%
  group_by(station_id) %>%
  summarize(mean_temp = mean(temp))
```

5. Which station had the coldest average temperature in January (hint: look up the arrange() function)
```{r}
# Mean for Jan for each station
df %>%
  filter(month == 1) %>%
  group_by(station_id) %>%
  summarize(mean_temp = mean(temp)) %>%
  arrange(mean_temp)
```