---
title: "Module 1 Assignment 3: Getting to Know your Home"
author: "Ellen Bledsoe"     # <- your name here!
date: "`r Sys.Date()`"      # <- uses the current date when rendered
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Assignment Description

### Purpose
The goal of this assignment is to get comfortable using the `tidyverse` with 
2-dimensional datasets and compare this process to using base R. 

### Task
Write R code using the `tidyverse` to successfully answer each question below.

### Criteria for Success
* Code is within the provided code chunks
* Code is commented with brief descriptions of what the code does
* Code chunks run without errors 
* Code produces the correct result
* This is the one time I _will_ take points off for not using `tidyverse`...

### Due Date


# Assignment Questions

For this final assignment for Module 1, you'll be working with another real-world
data set--a collection of data from climate stations scattered across Antarctica.

1. Load in the `tidyverse` package.
```{r load_tidyverse} 
# load packages
library(tidyverse)
```

2. Load in the data file (called aggregated_station_data.csv). It should already
be in your data folder. Save the data as an object called `weather`.

(Hint: remember to use the `tidyverse` function for reading in data! If it doesn't
pop up when you start typing, you probably forgot to run the code chunk above). 
```{r load_data}
# read in weather station data
weather <- read_csv("../data/aggregated_station_data.csv")
```

3. Use the `glimpse()` function to take a peek at the data. How many rows and 
columns are in the data? Write these answers in the code chunk as comments.
```{r glimpse_data} 
# explore the data
glimpse(weather)

# rows: 139, 144
# columns: 12
```

4. Write code to determine how many unique stations are in the dataset. 

(Hint: look up the help file for the `distinct()` function). 
```{r unique_stations}
# number of unique stations
weather %>% 
  distinct(station_id) %>% 
  count() # optional 

```

5. Write a chunk of code that calculates the mean temperature for the month of 
January for each station. 

Hint: think about using the pipe, `filter()`, `group_by()`, and `summarize()`. 
```{r mean_jan_temp}
weather %>% 
  filter(month == 1) %>% 
  group_by(station_id) %>% 
  summarize(mean_temp = mean(temp))

```

6. Write code to determine which station had the coldest average temperature in 
January.

(hint: look up the `arrange()` function)
```{r}
weather %>% 
  filter(month == 1) %>% 
  group_by(station_id) %>% 
  summarize(mean_temp = mean(temp)) %>% 
  arrange(mean_temp)
```