---
title: "Module 1 Assignment 3: Getting to Know your Home"
author: "Ellen Bledsoe"     # <- your name here!
date: "`r Sys.Date()`"      # <- uses the current date when rendered
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Description

### Purpose

The goal of this assignment is to get comfortable using the `tidyverse` with 2-dimensional data sets and compare this process to using base R.

### Task

Write R code using the `tidyverse` to successfully answer each question below.

### Criteria for Success

-   Code is within the provided code chunks
-   Code is commented with brief descriptions of what the code does
-   Code chunks run without errors
-   Code produces the correct result
-   This is the one time I *will* take points off for not using `tidyverse`...

### Due Date

Sept 15 at midnight MDT

# Assignment Questions

For this final assignment for Module 1, you'll be working with another real-world data set--a collection of data from climate stations scattered across Antarctica.

1.  In your own words, describe what the `tidyverse` is. Your answer should be between 1-3 sentences.

2.  Load in the `tidyverse` package.

```{r load_tidyverse}
# load packages
library(tidyverse)
```

3.  Load in the data file (called aggregated_station_data.csv). Save the data as an object called `weather`.

```{r load_data}
# read in weather station data
weather <- read_csv("../data/aggregated_station_data.csv")
```

4.  Take a look at the data in whichever way you would like (e.g., `glimpse()`, `slice()`, `str()`, `head()`, etc.). How many rows and columns are in the data? Type your answers below:

    rows: 139, 160

    columns: 12

5.  Create a data frame that includes temperatures which are above freezing (AKA greater than 0)

```{r paged.print=TRUE}
filter(weather, temp > 0)
```

6.  Create a new data frame called `temp` that includes *only* the following columns: year, day, month, temp, station_id.

```{r temp_df}
# note to graders: I accidently had the wrong code in the answer key
# some students might have year, day month columns, other might have what matches the answer key (hour, running_day, month)
# either is fine!
temp <- select(weather, year:month, temp, station_id)
temp
```

7.  Using the data frame you created in Q5 above (`temp`), add a new column to that data frame that converts the temperature column (currently in Celsius) to Fahrenheit. Call the new column `tempF`. (Hint: we did this in class--use that same equation)

```{r tempF}
temp %>% 
  mutate(tempF = temp*(9/5) + 32)
```

8.  In your own words (either bullet points or sentence form is fine), explain two benefits of using the pipe (`%>%`).

9.  Find the minimum temperature recorded for each month (in Celsius, the original column). (Hint: think about months first (split) and then temperature (apply). You will also want to remove all the NA values.)

```{r}
temp %>% 
  group_by(month) %>% 
  summarise(min_temp = min(temp, na.rm = TRUE))
```

10. Create a data frame with the mean temperature for the month of January for each station.

Some hints:

-   take note of how months are represented in the data
-   think about using the pipe, how we choose which rows we want, and how we split-apply-combine
-   remember to remove the NA values!

```{r mean_jan_temp}
weather %>% 
  filter(month == 1) %>% 
  group_by(station_id) %>% 
  summarize(mean_temp = mean(temp, na.rm = TRUE))
```

## Bonus! (up to 2 points)

Write code to determine how many unique stations are in the `weather` data set. (Hint: look up the help file for the `distinct()` and the `count()` functions).

```{r unique_stations}
# number of unique stations
weather %>% 
  distinct(station_id) %>% 
  count() 
```
