---
title: "Module 3 Assignment 1"
author: "Ellen Bledsoe" # <- your name here!
date: '2022-10-27'      # <- the date here!
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Details

### Purpose

The goal of this assignment is to assess your ability to compare means numerically, visually, and statistically

### Task

Write R code which produces the correct answers and correctly interpret the results of visualizations and statistical tests.

### Criteria for Success

-   Code is within the provided code chunks
-   Code is commented with brief descriptions of what the code does
-   Code chunks run without errors
-   Code produces the correct result
    -   Code that produces the correct answer will receive full credit
    -   Code attempts with logical direction will receive partial credit
-   Written answers address the questions in sufficient detail

### Due Date

April 4 at midnight MST

# Assignment Questions

In this assignment, we're going to explore another data set on wind turbines that generate a significant portion of the energy for us down here in Antarctica.

### Set-Up

Let's load the `tidyverse` and read in the data set. Call the data `turbines`.

```{r}

```

1.  Explore the data set, either through the environment or through code. Answer the following questions (2 point):

    a.  How many turbine makers are there?
    b.  What does each row of data represent?

    ```{r}
    # optional; only if you want space for coding
    ```

### Numeric

2.  Generate a summary of the data set that calculates the mean wind speed and mean power output for each wind turbine company. Save this as a data frame called `turbine_summary`. (2 point)

```{r}

```

### Visual

3.  Create a mutiple density plot for the power output variable. (3 points)
    -   be sure to have a density plot for each turbine producer; the color and the fill should be determined by the maker of the turbine
    -   add in vertical lines for the mean values in the same color as the turbine makers (remember to reference the correct data frame!)
    -   make sure the x-axis, y-axis, and legend labels are capitalized and easier to understand 
    -   use the `theme_classic()` function

```{r}

```

4.  Generate a box-and-whisker plot using `ggplot2` that compares the wind speed between different turbine makers (3 points).

    The plot should:

    -   have capitalized and more descriptive axis labels (hint: wind speed is measured in kilometers per hour---km/hr)
    -   show raw data points in addition to the boxes. The points should be jittered.
    -   use the `theme_classic()` function

```{r}

```

### Statistic

5.  Write a null hypothesis and an alternative hypothesis for the question we are asking and that we will be using statistics to answer. (2 points)

Power Output:
    **Null Hypothesis** (H~0~):\
    **Alternative Hypothesis** (H~A~):
    
Wind Speed:

-   **Null Hypothesis** (H~0~):\
-   **Alternative Hypothesis** (H~A~):

6.  Based on the mean values in the `turbine_summary` data frame and the plots you've created above, predict the outcome of each t-test that we will run: 1 comparing wind speeds, 1 comparing power outputs (graded for completion, not accuracy). Explain your reasoning (1-2 sentences for each t-test is fine). (2 points)

    *Answer:*

7.  Perform a t-test on the power output by turbine maker. (1 point)

```{r}

```

8.  In 2-3 sentences, interpret the output from question 7. Focus on what the p-value is in reference to the cutoff of 0.05, what that means, and whether that means we accept or reject the null hypothesis. (2 points)

    *Answer:*

9.  Perform another t-test, this time on the wind_speed variable by manufacturer. (1 point)

```{r}

```

10. In 2-3 sentences, interpret the output from question 9 (focus on the same ideas as question 8). (2 points)

    *Answer:*
