---
title: "Module 1: Introduction to R and RStudio"
author: "Keaton Wilson, Ellen Bledsoe"
date: "11/22/2019, revised July 2022"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to R/RStudio

## Learning Outcomes

-   Students will be able to describe what computer programming languages (code) are, with R as an example.
-   Students will be able to describe the role of RStudio in programming with R.
-   Students will be able to describe the utility of each panel in RStudio.
-   Students will be able to perform basic math functions in the R console.

## Check-in (5 minutes)

Let's all sign-up for RStudio Cloud!

## RStudio exploration (10 minutes)

Open up RStudio Cloud. Spend some time familiarizing yourself with the structure of what is in front of you. Individually, come up with a question or something you don't really understand about R/RStudio. Be ready to report out.

## Rstudio Mini-tour (10 minutes)

Perform a live-coding mini-tour of RStudio.\
1. What the different panels do\
2. Console versus a script/Rmarkdown file?

## Using R as a calculator

For example:

```{r}
3 + 5

888 - 111

4^2

```

### Let's practice in the console, too.

You can do basic math in the console (the bottom left part of the screen). The console only understands R code, so we don't need to use the `{r}` notation; we can type numbers and mathematical symbols. Try multiplying 5 and 3 (hint: \* means multiply) in the console.

### Math: Small Group Challenge

Run the following equation in the console. You'll need to be creative, dig into the help files and probably the web to figure out a few of the more complicated operations. Be ready to share your code.

The log (base 10) of the square root of pi, all of which is raised to the power of 12.

```{r}
log10(sqrt(pi))^12
```
