---
title: 'Module 1: Intro to Coding in R'
author: "Ellen Bledsoe"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 80
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to Coding

## Learning Outcomes

-   Students will be able to define the following terms:
    -   object
    -   assignment
    -   vector
    -   function
    -   data frame
-   Students will be able to run code line-by-line and as code chunks from an
    Rmarkdown file.
-   Students will be able to comment their code effectively.
-   Students will be able to write code assign values to variables and use these
    variables to perform various operations.
-   Students will be able to use help files to learn how to use functions.
-   Students will be able to recall and explain how functions operate, and the
    basic syntax around functions (arguments, auto-completion, parentheses).
-   Students will be able to differentiate different data classes in R.
-   Students will learn how to create their own data structures (vectors, data
    frames).

## Assigning Objects

Assignments are really key to almost everything we do in R. This is how we
create permanence in R. Anything can be saved to an object, and we do this with
the assignment operator, `<-`.

The short-cut for `<-` is `Alt + -` (or `Option + -` on a Mac)

```{r}

```

## 1-Dimensional Data: Vectors

We can also assign more complex group of elements of the same type to a
particular object. This is a basic data structure in R.

```{r}

```

## Data classes

There are a few main types in R, and they behave differently.

-   numeric (numbers)
    -   integer (no decimals allowed)
    -   double (decimals allowed---interchangeable with numeric)
-   character (letters or mixture)
-   logical (True or False; T or F)
-   factors (best used for data that need to be in a specific order; levels
    indicate the order)

```{r}

```

Vectors have to contain elements that are all of the same class.

```{r, eval = FALSE}

```

### Sub-setting Vectors

Sometimes we want to pull out and work with specific values from a vector. This
is called sub-setting (taking a smaller set of the original)

```{r}

```

## Functions

Functions are pre-written bits of codes that perform specific tasks for us.

Functions are always followed by parentheses. Anything you type into the
parentheses are called arguments.

```{r}

```

To get more information about a function, use the `help()` function or
`?name_of_function`.

```{r}

```

For examples, we can use a function called `class()` to figure out the data type
of a vector.

```{r}

```

### Group Challenge

Let's practice! Write a few lines of code that do the following:

-   create a vector with a number from 6 to 1
-   assign the vector to an object named `vec`
-   subset `vec` to include the last 3 numbers (should include 3, 2, 1)
-   find the sum of the numbers (hint: use the `sum()` function)

```{r}

```

Already finished? See if you can condense your code down any further or turn
around and help out a neighbor.

## 2-Dimensional Data: Data Frames

Most of the data you will encounter is two-dimensional, i.e., it has columns and
rows. Its structure resembles a spreadsheet. R is really good with these types
of data.

-   **rows** go side-to-side
-   **columns** go up-and-down

Data frames are made up of multiple vectors. Each vector becomes a column.

```{r}

```

### Sub-setting Data Frames

Because data frames are two-dimensional, we can subset data in different ways.
We can select specific columns, specific rows, or filter rows by values.

R always takes information for the row first, then the column.

```{r}

```

### Discussion point

Plants is a simple data set, but let's come up with some questions.

```{r}

```

### Group Challenge (5 min)

As a group, find the standard deviation (`sd()`)of the height of plants treated
with nitrogen and those not treated with nitrogen. Which group has the larger
standard deviation?

```{r}

```

Come up with a definition of standard deviation, use the help file to find out
how the `sd()` function works, and be prepared to show the code you used.

*Definition*:

## Helpful Functions

Below are some functions that I often find very helpful when working with
vectors and data frames:

-   `str()`
-   `head()` and `tail()`
-   `length()`
-   `ncol()` and `nrow()`
-   `names()`

```{r}

```

# Rmarkdown and Code Chunks (10 minutes)

Rmarkdown (.Rmd) is a file format that lets us incorporate text and code into
one document seamlessly. In fact, it is the file format for this document!

-   For writing text, you can type as you would normally
-   Code chunks are a bit different.
    -   Type your R code in the lines under the `{r}`.

    -   To include text in them, you will need to put a `#` in front. R will not
        read anything after the `#` as code. Chunks look like this:

```{r}

```

A quick shortcut for adding a code chunk is `Ctrl + Atl + i` (`Cmd + Opt + i` on
a Mac). Alternatively, you can go to Code \> Insert Chunk.

To run a chunk of code, click the green arrow on the far right side of the
chunk. You can also run one or a few lines of code at a time by having your
cursor on the line or highlighting multiple lines and hitting `Ctrl` + `Enter`
(or `Cmd` + `Enter` on a Mac).

Let's work with an example code chunk:

```{r}

```

If you run code in the console (below) rather than the code chunk in a .Rmd
file, you don't need to add the `{r}` to tell it that you are typing R code. The
console only understand R code, so we don't need to tell it!

### Group challenge

Insert a new code chunk in this section. In the code chunk, calculate the
average (mean) circumference of trees in the orange data set. Save the mean as
an object called "avg_circ."

If you're already finished, try this challenge. Find the average circumference
of trees aged 484 days.
